---
title: "Homework 6 solutions"
author: "Jeff Goldsmith"
output: github_document
---


```{r, load_libraries, message = FALSE, warning=FALSE}
library(tidyverse)
library(ggridges)
```



## Problem 2
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```
### Bootstrap $\hat{r}^2$ value
```{r bootstrap_r_squared}
bootstrap_r_squared = weather_df |> 
  modelr::bootstrap(n=5000) |> 
  mutate (
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data=df)),
    results = map(models,  broom::glance) 
  ) |> 
  select(-strap, -models) |> 
  unnest(results) |> 
  select(r.squared)

```

#### Ploting distribution of $\hat{r}^2$ values
```{r}
ggplot(aes(x = r.squared), data=bootstrap_r_squared) + geom_density()
```
The computed $\hat{r}^2$ is high with an average of `r mean(boostrap_r_squared$r.squared)`, which reflects a significant proportion of the variability in the dependent variable `tmax` can be explained by the independent variables `tmin` and `prcp`. The min of the bootstrap values of $\hat{r}^2$ is `r min(boostrap_r_squared$r.squared)` and the max is `r max(boostrap_r_squared$r.squared)`

#### 95% Confidence Interval of $\hat{r}^2$ values
```{r}
bootstrap_r_squared  |> 
  summarize(
    ci_lower = quantile(r.squared, 0.025), 
    ci_upper = quantile(r.squared, 0.975))
```

